---
title: "Body measurements"
author: "Amelia Ritger"
date: "8/25/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
library(janitor)
library(gtools)
library(lubridate)
library(ggpmisc) #use for adding regression line equation to ggplot
```

## Load data
```{r}
data <- read_csv(here("data", "Measurements.csv")) %>% 
  clean_names()
```

## Tidy data
```{r}
# convert NAs to 9999 for minimum value calculation next
data_no_na <- data %>%
  replace_na(list(dry_mass_plus_vial_g_day_1="9999")) %>%
  replace_na(list(dry_mass_plus_vial_g_day_2="9999")) %>%
  replace_na(list(dry_mass_plus_vial_g_day_3="9999")) %>%
  replace_na(list(dry_mass_plus_vial_g_day_4="9999")) %>%
  replace_na(list(dry_mass_plus_vial_g_day_5="9999")) %>%
  replace_na(list(dry_mass_plus_vial_g_day_6="9999")) %>%
  replace_na(list(dry_mass_plus_vial_g_day_7="9999"))

# Take the measurements, do algebra
measure <- data_no_na %>%
  mutate(oral_disk_mm = (oral_disk_diameter_mm_a + oral_disk_diameter_mm_b)/2) %>%
  mutate(basal_disk_mm = (basal_disk_diameter_mm_a + basal_disk_diameter_mm_b)/2) %>%
  mutate(wet_mass_g = wet_mass_plus_vial_g-vial_mass_g) %>%
  mutate(min_dry_mass_plus_vial_g = pmin(dry_mass_plus_vial_g_day_2, dry_mass_plus_vial_g_day_3,dry_mass_plus_vial_g_day_4,dry_mass_plus_vial_g_day_5, dry_mass_plus_vial_g_day_6, dry_mass_plus_vial_g_day_7)) %>%
  mutate(dry_mass_g = as.numeric(min_dry_mass_plus_vial_g)-vial_mass_g) %>%
  select(number, date, id, oral_disk_mm, basal_disk_mm, mouth_length_mm, vial_mass_g, min_dry_mass_plus_vial_g, wet_mass_g, dry_mass_g) %>% 
  filter(is.na(dry_mass_g)==FALSE) %>% 
  mutate(dry_mass_g=ifelse(dry_mass_g <= 0, 0.00001, dry_mass_g)) #make negative or zero values (samples 85, 160) very small greater than zero values

measure_dry <- measure %>% 
  filter(dry_mass_g>0) #remove zero values

```

Univariate analysis
```{r}
#basal disk diameter
hist(measure$basal_disk_mm)

ggplot(measure, aes(sample=basal_disk_mm)) +
  stat_qq() +
  stat_qq_line()

#dry mass
hist(measure$dry_mass_g)

ggplot(measure, aes(sample=dry_mass_g)) +
  stat_qq() +
  stat_qq_line()
```

Well those variables are quite skewed!

### Transform those bad boys
```{r}
measure <- measure %>% 
  mutate(basal_log = log10(basal_disk_mm),
         dry_log = log10(dry_mass_g))

#log(basal disk diameter)
hist(measure$basal_log)

ggplot(measure, aes(sample=basal_log)) +
  stat_qq() +
  stat_qq_line()

#log(dry mass)
hist(measure$dry_log)

ggplot(measure, aes(sample=dry_log)) +
  stat_qq() +
  stat_qq_line()
```

That looks better!

# Plot some stuff
```{r}
formula1 <- y ~ x
formula2 <- y ~ poly(x,2)
formula3 <- y ~ poly(x,3)

# oral disk ~ mouth length, dry mass/wet mass ~ oral disk/mouth length
ggplot(measure, aes(x=basal_disk_mm, y=dry_mass_g)) +
  geom_point() +
  #geom_point(aes(color=id)) + # plot points colored by ID
  geom_smooth(method="lm", formula= (y ~ x), color="lightgreen", fill="lightgreen", linetype="dashed") + 
  geom_smooth(method="lm", formula= (y ~ poly(x,2)), color="blue", fill="blue", linetype="dashed") +
  geom_smooth(method="lm", formula= (y ~ poly(x,3)), color="red", fill="red", linetype="dashed") +
  #geom_smooth(method = "lm", se = FALSE, formula = y ~ exp(x))
  labs(x="Oral Disk Diameter (mm)", y="Wet mass (g)") +
  theme_bw() +
  stat_poly_eq(formula = formula1, 
                aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")), 
               label.y = 0.95, label.x = 0.025,
               parse = TRUE) +
  stat_poly_eq(formula = formula2, 
                aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")),
                label.y = 0.875, label.x = 0.025,
                parse = TRUE) +
  stat_poly_eq(formula = formula3, 
                aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")), 
               label.y = 0.8, label.x = 0.025,
                parse = TRUE)
```

## Plot log-transformed variables just for fun
```{r}
ggplot(measure, aes(x=basal_log, y=dry_log)) +
  geom_point() +
  #geom_point(aes(color=id)) + # plot points colored by ID
  geom_smooth(method="lm", formula= (y ~ x), color="lightgreen", fill="lightgreen", linetype="dashed") +
  labs(x="log(Oral Disk Diameter) (mm)", y="log(Wet mass) (g)") +
  theme_bw() +
  stat_poly_eq(formula = formula1, 
                aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")), 
               label.y = 0.95, label.x = 0.025,
               parse = TRUE)
```

## Run the regression
```{r}
#first, remove missing values (samples where I didn't measure basal disk diameter)
measure_na <- measure %>% 
  filter(!is.na(basal_disk_mm))
```

log-log transformed data
```{r}
model1 <- lm(measure_na$dry_log ~ measure_na$basal_log)

summary(model1)
AIC(model1)
```

2nd degree polynomial
```{r}
model2 <- lm(measure_na$dry_mass_g ~ poly(measure_na$basal_disk_mm,2))

summary(model2)
AIC(model2)
```

3rd degree polynomial
```{r}
model3 <- lm(measure_na$dry_mass_g ~ poly(measure_na$basal_disk_mm,3))

summary(model3)
AIC(model3)
```

Hmm... it actually looks like I don't need third degree polynomial? BUT ALSO WHY IS AIC SO LARGE (negative doesn't matter) WHEN A POLYNOMIAL IS USED GAH I don't think AIC is a good indicator for model selection here

Some of these points are looking funny, what if I exclude everything measured pre-2020?
```{r}
measure_2020 <- measure %>% 
  mutate(year=lubridate::year(as.Date(date, format = "%m/%d/%Y"))) %>% 
  filter(year=="2020")

ggplot(measure_2020, aes(x=basal_log, y=dry_log)) +
  geom_point() +
  #geom_point(aes(color=id)) + # plot points colored by ID
  geom_smooth(method="lm", formula= (y ~ x), color="lightgreen", fill="lightgreen", linetype="dashed") +
  labs(x="log(Oral Disk Diameter) (mm)", y="log(Wet mass) (g)") +
  theme_bw() +
  stat_poly_eq(formula = formula1, 
               aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")), 
               label.y = 0.95, label.x = 0.025,
               parse = TRUE)
```




